---
title: "Day-08.09 - ESS 330 Daily Exercises"
author: "Yazeed Aljohani"
date: "`r Sys.Date()`"
format: html
editor: visual
---

```{r}

library(visdat)
library(broom)
library(ggplot2)
```

# **1.. Loading and Exploring the Dataset**

```{r}
# Loading the airquality dataset
data("airquality")

```

```{r}
# Using help function to explore the dataset
?airquality

```

# 2.Checking for Missing Values and Cleaning the Data

```{r}
# Visualizing missing values
vis_dat(airquality)
```

```{r}
# Removing missing values
airquality_cleaned <- na.omit(airquality)

```

# 3.Fitting a Linear Model to Predict Ozone from Temperature

```{r}
# Fitting a linear model to predict Ozone from Temp
model <- lm(Ozone ~ Temp, data = airquality_cleaned)
```

# 4.**Viewing and Interpreting the Model Summary**

```{r}
# Viewing model summary
summary(model)
```

-   **R² = 0.488** means that **48.8% of the variation in Ozone is explained by Temp**.

-   **Adjusted R² (0.4833)** accounts for the sample size and number of predictors.

### **Assessing Model Validity**

1.  **Significance of Temp**:

    -   The **t-value (10.192)** is large.

    -   The **p-value (\<2e-16)** confirms that Temp is a strong predictor.

2.  **Residual standard error: 23.92 on 109 degrees of freedom:**

-   This indicates that our predictions have an average error of **±23.92 Ozone units**.

-   Large residuals suggest that additional factors (e.g., Wind, Solar.R) might improve predictions.

**F-statistic (103.9, p \< 2.2e-16)**:

-   The model as a whole is **statistically significant**.

-   This means that Temp is not a random predictor but has a real effect on Ozone.

### **4. Explain R-squared**

R² measures how well the predictor (Temp) explains the variability in Ozone.

**Interpretation**:

> **"48.8% of the variation in Ozone levels can be explained by Temperature."**

-   The remaining **51.2%** is due to **other unknown factors**.

# 5. Predicting Ozone Using the Cleaned Data

```{r}
# Predictinf Ozone using the cleaned data
predictions <- augment(model, newdata = airquality_cleaned)
```

```{r}
# Ensuring column names match for correlation calculation
colnames(predictions)[colnames(predictions) == ".fitted"] <- "Predicted_Ozone"
```

# 6. Plotting Actual vs. Predicted Ozone

```{r}
# Ploting actual vs predicted Ozone
ggplot(predictions, aes(x = Ozone, y = Predicted_Ozone)) +
  geom_point(alpha = 0.7) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Ozone",
    x = "Actual Ozone",
    y = "Predicted Ozone",
    subtitle = paste("Correlation:", round(cor(predictions$Ozone, predictions$Predicted_Ozone, use = "complete.obs"), 2))
  ) +
  theme_minimal()
```

#### **Explanation of the Plot**

1.  **Scatter Plot**:

    -   Each point represents an observation.

    -   If the model were perfect, all points would lie **on the red line**.

2.  **Red 1:1 Line (`geom_abline`)**:

    -   Represents where **actual = predicted**.

    -   The closer the points are to this line, the **better the model's predictions**.

3.  **Correlation in Subtitle**:

    -   Displays `cor(Ozone, Predicted_Ozone)`, indicating how well predictions align with actual values.
